---
---
title: "Analysis of Statenisland_data""
output: html
---

##Download and Analyse the dataset
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/ASHMOHAN 1/Desktop/Data_Science/SMU_Term1_Doing_Data_science/Project1/dds_statenisland_assignment")

# So, save the file as a csv and use read.csv instead
statenisland_data <- read.csv("Data/rollingsales_statenisland.csv", header = TRUE, sep = ",")

## Check the dataset that shows the rolling house sales in statenisland.
head(statenisland_data)
summary(statenisland_data)
str(statenisland_data)
names(statenisland_data)
```

##Cleaning Data downloaded
Format some columns like square feet, year built and sale price in the dataset. 

```{r Cleaning Data}
statenisland_data$LAND.SQUARE.FEET <- as.numeric(gsub("[^[:digit:]]", "", statenisland_data$LAND.SQUARE.FEET))
statenisland_data$GROSS.SQUARE.FEET <- as.numeric(gsub("[^[:digit:]]", "", statenisland_data$GROSS.SQUARE.FEET))
statenisland_data$YEAR.BUILT <- as.numeric(as.character(statenisland_data$YEAR.BUILT))
statenisland_data$SALE.PRICE <- as.numeric(gsub("[^[:digit:]]", "", statenisland_data$SALE.PRICE))
```

#Subset the dataset and take required fields, correlate between the square feet of land and the sale price.
```{r Subsetting Data}
st.subset <- subset(x = statenisland_data, 
  ((statenisland_data$LAND.SQUARE.FEET > 1000) &  
     (statenisland_data$SALE.PRICE > 1000)), select = c(15,17,20))
str(st.subset)
cor(st.subset$LAND.SQUARE.FEET, st.subset$SALE.PRICE)
```

##Plot showing relation between year built and square feet. we see lots of houses are built after the year 1980. Later a table of the year built is done to double check if this is true.
```{r Plotting Data}
st.subset$SALE.PRICE <- (st.subset$SALE.PRICE/1000)
st.subset2 <- subset(x = statenisland_data, (statenisland_data$YEAR.BUILT > 1900 & 
  statenisland_data$LAND.SQUARE.FEET > 1000 & statenisland_data$LAND.SQUARE.FEET < 20000), select = c(15, 17))

str(st.subset2)

plot(st.subset2$YEAR.BUILT, st.subset2$LAND.SQUARE.FEET, main = "Square Feet Vs year Built", xlab= "Year Built", ylab = "Square Feet ")
table(st.subset2$YEAR.BUILT)
```


## Another plot -  let's look at family homes of 2000 - 2500 square feet and plot the prices
```{r Plotting mid size family homes and prices}

st.mid_size_home <- statenisland_data$LAND.SQUARE.FEET[which(statenisland_data$LAND.SQUARE.FEET > 2000 & statenisland_data$LAND.SQUARE.FEET < 2500)]
st.mid_size_home_value <- statenisland_data$SALE.PRICE[which(statenisland_data$LAND.SQUARE.FEET > 2000 & statenisland_data$LAND.SQUARE.FEET < 2500)]
hist(st.mid_size_home)
plot(st.mid_size_home, st.mid_size_home_value,main="Mid Size House Prices", xlab="Sale Price in dollars", ylab="Square feet", pch=19)
```
